The Chemistry Development Kit (CDK). 3. Atom typing, Rendering, Molecular Formula, and Substructure Searching

### 2016-07-16 article.tex
Started writing again; will try to send around a full draft around by the end of this week

  This CDK version has isolated a new atom typing framework, removing the perception of
  atom types from various algorithms, allowing the perception to be tested separately.
  The new code defines the atom types using a list that specifies for each type
  the element symbol, hybridization, formal charge, number of lone pairs, and
  an enumeration of the bond orders. This list of properties captures the
  information needed for the various algorithms in the CDK. For example,
  hybridization information can be used in certain aromaticity models (see later),
  and the lone pair information is needed for resonance structure calculation
  needed, for example, for Gasteiger \pi-charges.

@@

  The structure diagram layout ... %FIXME: John, do you want this in?

@@

Another recent change is the choice of build system: we have moved away from
Ant to Maven to deal with dependencies and compile the library.

% FIXME: write more about this, and provide info about Maven Central

@@

INVITE ARVID

@@

>>>>

  This CDK version has isolated a new atom typing framework, removing the perception of
  atom types from various algorithms, allowing the perception to be tested separately.
  The new code defines the atom types using a list that specifies for each type
  the element symbol, hybridization, formal charge, number of lone pairs, and
  an enumeration of the bond orders. This list of properties captures the
  information needed for the various algorithms in the CDK. For example,
  hybridization information can be used in certain aromaticity models (see later),
  and the lone pair information is needed for resonance structure calculation
  needed, for example, for Gasteiger $\pi$-charges.

@@

  The structure diagram layout has been rewritten and the new code solves a
  number of long standing issues. Particularly, collission detection has been
  greatly improved. Figure~\ref{fig:sdg} shows a ...

@@

Another recent change is the choice of build system: we have moved away from
Ant to Maven to deal with dependencies and compile and test the library.
To keep the original idea of CDK modules, this move required to split up the
source code in multiple source folders. Fortunately, modern integrated
development environments have no trouble handling this, removing the original
argument to have everything in one source folder.

On the other hand, for many modules, the test code is now more closely closely
linked to the code being tested: both reside in the same folder, though we
adhere to the Maven custom to have src/main/java and a src/test/java folders.
For a few modules, however, this solution introduces circular dependencies, in
which case a separate Maven module is created for the tests.

The Maven packages for the CDK are available from Maven Central, which makes it
easy for other projects to use. The full library can be included in other
software by depending on the cdk artifact (\url{http://mvnrepository.com/artifact/org.openscience.cdk/cdk})
but dependencies can also be defined on individual CDK modules.

@@

OSGi bundles are available for the CDK too, which are used by Bioclipse and
KNIME.

% ping Arvid

@@

\subsection*{Figure 3 - The improved structure diagram generation has improved
code to solve overlap. The original SDG code used general heuristics (left) and the
OverlapResolver would fine tune the layout to ensure atoms would not be placed
at the same location (middle). The new SDG algorithm is able to
make more rigorous changes, making the final output must more pleasing (right).}

\includegraphics[width=\textwidth]{sdg.png}
